---
title: "4 Environmental Data"
author: "Gemma Clucas"
date: "2/23/2021"
output: github_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(raster)
library(tidyverse)
library(rgdal)
library(dismo)
library(marmap)
library(ggfortify)
library(metR)
library(rgeos)
library(geosphere)

select <- dplyr::select
```

## Load tracking data from chick-rearing period
Read it in and find the extent of the area covered - this will be the study area.
```{r}
All <- read.csv("Chick-rearing_trips/All_chick-rearing_trips.csv", stringsAsFactors = FALSE)

# Make it spatial
coordinates(All) <- ~LON + LAT
proj4string(All) <- CRS("+proj=laea +lon_0=-26 +lat_0=-58 +units=m")
# Reproject to WGS84
All <- spTransform(All, CRS = CRS("+proj=longlat +ellps=WGS84"))

# Find the extent of the chick-rearing tracks
extent(All)
```


## Create a raster of the study area with land masked

This just creates a raster with all values set to 1 across the study area.
```{r}
# Read in bathymetry raster and crop to extent
SSI_bath_WGS84 <- raster("ssi_geotif/full_ssi18a.tif") %>% 
  projectRaster(., crs=crs("+init=epsg:4326")) %>% 
  crop(., c(-27.96598, -24.623, -58.41806, -57.28708))

# Set all values to 1
x <- SSI_bath_WGS84
values(x) <- 1
```


Crop out land from the study area raster using ```mask()```.
```{r}
# Read in shapefile for land
SSI_WGS84 <- readOGR("Seamask.shp") %>% 
  crop(., c(450000, 1095192, -795043.9, -100000)) %>% 
  spTransform(., crs("+init=epsg:4326"))

# Use mask() to cut out land from the raster
mask<-mask(x, SSI_WGS84, inverse=F)

# Plot to check
plot(mask)
```

**Should I be doing all this in LAEA?**

## Sample the raster layer to create 'background' points

This just creates a random sample of 'background' points across the study area raster. It takes a little while for this to run, so I have saved the dataframe as a CSV and just read it in each time for efficiency.

```{r}
# birds<-unique(All$Ptt)
# background<-data.frame()
# i<-birds[1]
# #RUN THROUGH EACH BIRD IN TURN
# for(i in birds){
#   bird<-All[All$Ptt==i,] #subset 1 bird
#   back <- randomPoints(mask, n = (nrow(bird)) ,prob=F) # this was originally 3*nrow
#   dat<-as.data.frame(back)
#   head(dat)
#   names(dat)<-c("Lon","Lat")
#   dat$bird<-i
#   background<-rbind(background,dat)
# }
# head(background)
# 
# write.csv(background, "background_points.csv", sep = "", row.names = FALSE)

background <- read.csv("background_points.csv", stringsAsFactors = FALSE)

# Plot the points to check
as.data.frame(mask, xy = TRUE) %>%
  ggplot() +
  geom_raster(aes(x=x, y=y)) +
  geom_point(data = background, aes(x=Lon, y=Lat)) +
  ylab("Latitude") +
  xlab("Longitude")

```

## Bathymetry
I've already read in the bathymetry raster, so I'm just masking land and plotting.
```{r}
# Use mask() to cut out land from the bathymetry raster
# values(SSI_bath_WGS84) <- 1
bathy_mask<-mask(SSI_bath_WGS84, SSI_WGS84, inverse=F)

# Make a plot using the marmap package
dat <- marmap::as.bathy(bathy_mask)
autoplot(dat, geom=c("raster", "contour"), coast = FALSE, colour="white", size=0.1) + 
  scale_fill_gradient(low = "steelblue4", high = "lightblue") +
  ylab("Latitude") +
  xlab("Longitude") +
  labs(fill = "Depth") 
  # xlim(c(-26.5, -26.35)) +
  # ylim(c(-57.875, -57.79))
```


## Distance from the colony
```{r}
colony_lat<- -57.808 
colony_lon<- -26.404

# Find the colony cell in the study area raster
j <- cellFromXY(mask, cbind(colony_lon, colony_lat))
# Change the value of the cell where the colony is to 2 (all the other cells are 1)
mask[j]<-2 

# Check it worked
# table(values(mask)) 

# Create a distance raster from the colony
dist<-gridDistance(mask, origin=2,omit=NA)
plot(dist)
```

If I do everything in LAEA, how do I convert the lon/lat of the colony to LAEA?

## Distance to shelf break

Firstly, at what depth does the shelf break occur?
```{r}
# plotting using metR package - can't figure out how to make it label all contour lines
autoplot(dat, geom=c("raster"), coast = FALSE, colour="white", size=0.1) + 
  scale_fill_gradient(low = "steelblue4", high = "lightblue") +
  geom_contour(data = as.data.frame(bathy_mask, xy = TRUE), 
               aes(x=x, y=y, z = full_ssi18a), 
               colour = "white",
               binwidth = 100) +
  geom_text_contour(data = as.data.frame(bathy_mask, xy = TRUE), 
                    aes(z = full_ssi18a), 
                    stroke = 0.2, 
                    skip = 0,
                    label.placement = label_placement_flattest(2)) +
  ylab("Latitude") +
  xlab("Longitude") +
  labs(colour = "Depth") +
  xlim(c(-26.7, -26.2)) +
  ylim(c(-57.9, -57.6))
```

It seems like the shelf break occurs at around 500 - 1000m depth.

So I need to make a raster that includes the distance to the shelf break from all points in the study area. I'm going to use the 500m depth contour.



```{r}
# first draw contours every 100m
contours <- rasterToContour(bathy_mask, nlevels = 46)
plot(bathy_mask)
plot(contours, add=TRUE)

# select just the 500m depth contour
contour500 <- contours[contours@data$level == -500, ]

# check the contour and the study area raster have the same projection
# crs(contour500)
# crs(mask)
# yes they do

# Make the SpatialLinesDataFrame (contour500) into a raster with the same characteristics as the study area raster (mask)
# Set the contour pixels to 1, all others should be NA
r500 <- rasterize(contour500, mask, field = 1, background = NA)

# Calculate the distance to the nearest non-NA cell
shelfdist <- raster::distance(r500)

# Mask the land
shelfdist <- mask(shelfdist, SSI_WGS84, inverse=F)
plot(shelfdist)

```

## Bearing to the shelf break

```{r}
# get the values from the distance to colony raster along the 500m depth contour and find the minimum value
# i.e. the minimum distance of the depth contour from the colony
min500dist <- raster::extract(dist,contour500) %>% 
  unlist() %>% 
  min()

# Change the raster into a dataframe, so that we can extract the coordinates of this minimum point
as.data.frame(rasterToPoints(dist)) %>% 
  dplyr::filter(layer == min500dist) %>% 
  select(x,y)

# So these are the coordinates of the nearest bit of shelf break
shelf_lon <- -26.36173
shelf_lat <- -57.82149

bearing <- geosphere::bearingRhumb(c(colony_lon, colony_lat), c(shelf_lon, shelf_lat))

```

So the bearing to the nearest bit of shelf is `r bearing` but this direction is actually blocked by land... what do I do about that? 




### Messing around with KDEs and density plots
This should go in a separate script really and is not finished. 
Can use this to formally calculate a KDE and plot
```{r}
library(spatialEco)
kde <- sp.kde(All, bw = 0.01, 
              nr = 50,
              nc = 50,
              standardize = TRUE)

ggplot(data = as.data.frame(kde, xy = TRUE), aes(x=x, y=y)) +
  geom_contour_fill(aes(z = kde)) +
  #scale_fill_gradient(low = "yellow", high = "red") +
  scale_alpha(range = c(0.00, 0.5), guide = FALSE) 
```

Or you can just plot the density directly from the ```SpatialPointsDataFrame``` using ```stat_density2d()```.
```{r}
ggplot(data = as.data.frame(All), aes(x = LON, y = LAT)) +
  stat_density2d(aes(alpha=..level..), geom = "polygon", fill = "orange") +
  scale_alpha(range=c(0.1,0.75),guide=FALSE)

```

Plotting on top of bathymetry map
```{r}
autoplot(dat, geom=c("raster", "contour"), coast = FALSE, colour="white", size=0.1) + 
  scale_fill_gradient(low = "steelblue4", high = "lightblue") +
  ylab("Latitude") +
  xlab("Longitude") +
  labs(fill = "Depth") +
  stat_density2d(data = as.data.frame(All),
                 aes(x = LON, y = LAT, alpha = ..level..), 
                 geom = "polygon", 
                 fill = "red") +
  scale_alpha(range=c(0.2,0.9),guide=FALSE)
```

## Other things
Look at google earth engine for SST/chlorophylA: https://developers.google.com/earth-engine/datasets/catalog/NASA_OCEANDATA_MODIS-Terra_L3SMI

