---
title: "5_GAMs"
author: "Gemma Clucas"
date: "3/24/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mgcv)
library(pROC)
library(tidyverse)
```

Read in my data
```{r}
data <- read.csv(file = "PresBackgroundLocationsWithEnvironmentalVariables.csv", stringsAsFactors = FALSE)

birds <- unique(data$Ptt)

names(data)

#birds <- "196698"
```


Run the GAMs varying the number of knots from 3 to 5. This holds out the real and background data for each Ptt in turn, and then calculates the mean AUC, sensitivity, and specificty from all 20 runs.

I can't seem to get the formula for the GAM to take anything but the actual variable name, so I am not able to run through each variable automatically at the moment.
```{r}
var <- "chlorA"
k <- c(3,4,5)
# k <- c(3)

for (i in k) {
  for (j in birds){
    TRAIN <- data[data$Ptt!=j,]  # use all groups except 1 in training data
    TEST<- data[data$Ptt==j,]  # keep the other group for testing data
    # run gam with one variable, k is number of knots
    GAM <- gam(pres ~ s(chlorA, k = i), data=TRAIN, bs="cs", family=binomial, select=TRUE, method='GCV.Cp')  
    TEST$GAM_pred <- as.numeric(predict(GAM, type="response", newdata=TEST)) # predict into the test data frame
    roc1 <- roc(TEST$pres, TEST$GAM_pred) # create a roc curve from the test data
    AUCEVAL <- as.numeric(roc1[9]) # get the auc value
    co1 <- pROC::coords(roc1, x = "best", best.method = "closest.topleft", ret=c("specificity","sensitivity")) #also get specificity and sensitivity
    spec1 <- as.numeric(co1[1])
    sen1 <- as.numeric(co1[2])
    eva <- as.data.frame(rbind(AUCEVAL,spec1,sen1))
    names(eva) <- j
    eva
    if(exists("allevals")){
      allevals <- cbind(allevals, eva)
    }else {
      allevals <- eva
    }
  }
  # calculate mean AUC, sensitiviy, and specificity
  allevals$mean <- rowMeans(allevals)
  # save as a csv
  write.csv(allevals, file = paste("GAMs/", var, "_k", i, ".csv", sep = "" ))
  # save just the mean, with the variable names to an object
  assign(paste(var, i, sep = "_k"), 
       as_tibble(cbind(group = names(allevals), t(allevals))) %>%
         filter(group == "mean") %>% 
         select(-group) %>% 
         mutate(var = var, k = i) )
  rm(allevals)
}
```

Trying to automate the covariates
```{r}
var <- c("depth", "colonydist")
knots <- c(3,4,5)


for (l in var) {
  for (i in knots) {
    for (j in birds){
      TRAIN <- data[data$Ptt!=j,]  # use all groups except 1 in training data
      TEST<- data[data$Ptt==j,]  # keep the other group for testing data
      # run gam with one variable, k is number of knots
      form <- as.formula(paste("pres ~ s(", l, ", k = i)"))
      GAM <- gam(form, data=TRAIN, bs="cs", family=binomial, select=TRUE, method='GCV.Cp')  
      TEST$GAM_pred <- as.numeric(predict(GAM, type="response", newdata=TEST)) # predict into the test data frame
      roc1 <- roc(TEST$pres, TEST$GAM_pred) # create a roc curve from the test data
      AUCEVAL <- as.numeric(roc1[9]) # get the auc value
      co1 <- pROC::coords(roc1, x = "best", best.method = "closest.topleft", ret=c("specificity","sensitivity")) #also get specificity and sensitivity
      spec1 <- as.numeric(co1[1])
      sen1 <- as.numeric(co1[2])
      eva <- as.data.frame(rbind(AUCEVAL,spec1,sen1))
      names(eva) <- j
      eva
      if(exists("allevals")){
        allevals <- cbind(allevals, eva)
      }else {
        allevals <- eva
      }
    }
    # calculate mean AUC, sensitiviy, and specificity
    allevals$mean <- rowMeans(allevals)
    # save as a csv
    write.csv(allevals, file = paste("GAMs/", l, "_k", i, ".csv", sep = "" ))
    # save just the mean, with the variable names to an object
    assign(paste(l, i, sep = "_k"), 
           as_tibble(cbind(group = names(allevals), t(allevals))) %>%
             filter(group == "mean") %>% 
             select(-group) %>% 
             mutate(var = l, k = i) )
    rm(allevals)
  }
}
```




Combine the outputs from all the different GAMs with just one covariate, to find the one with the best AUC. 
```{r}
  
dplyr::bind_rows(colonydist_k3, colonydist_k4, colonydist_k5,
                 depth_k3, depth_k4, depth_k5) %>% 
                # ,
                #  slope_k3, slope_k4, slope_k5,
                #  SST_k3, SST_k4, SST_k5,
                #  Height_k3, Height_k4, Height_k5,
                #  chlorA_k3, chlorA_k4, chlorA_k5,
                #  NorthVelocity_k3, NorthVelocity_k4, NorthVelocity_k5,
                #  EastVelocity_k3, EastVelocity_k4, EastVelocity_k5) %>%
  select(var, k, AUCEVAL, spec1, sen1) %>% 
  rename(covariate = var, knots = k,
         AUC = AUCEVAL, Sensitivity = sen1, Specificity = spec1) %>% 
  arrange(desc(AUC))
```

So colonydist is the best predictor. Run models with colonydist plus the other covariates next.
Also try using different methods for "best" threshold determination.


```{r}
var <- "depth"
k <- c(3,4,5)
# k <- c(3)

for (i in k) {
  for (j in birds){
    TRAIN <- data[data$Ptt!=j,]  # use all groups except 1 in training data
    TEST<- data[data$Ptt==j,]  # keep the other group for testing data
    # run gam with one colonydist, knots = 3, and vary the other covariates and number of knots
    GAM <- gam(pres ~ s(colonydist, k = 3) + s(depth, k = i), data=TRAIN, bs="cs", family=binomial, select=TRUE, method='GCV.Cp')  
    TEST$GAM_pred <- as.numeric(predict(GAM, type="response", newdata=TEST)) # predict into the test data frame
    roc2 <- roc(TEST$pres, TEST$GAM_pred) # create a roc curve from the test data
    AUCEVAL2 <- as.numeric(roc2[9]) # get the auc value
    co2 <- pROC::coords(roc2, x = "best", best.method = "closest.topleft", ret=c("specificity","sensitivity")) #also get specificity and sensitivity
    spec2 <- as.numeric(co2[1])
    sen2 <- as.numeric(co2[2])
    eva2 <- as.data.frame(rbind(AUCEVAL2,spec2,sen2))
    names(eva2) <- j
    eva2
    if(exists("allevals2")){
      allevals2 <- cbind(allevals2, eva2)
    }else {
      allevals2 <- eva2
    }
  }
  # calculate mean AUC, sensitiviy, and specificity
  allevals2$mean <- rowMeans(allevals2)
  # save as a csv
  write.csv(allevals2, file = paste("GAMs/colonydist+", var, "_k", i, ".csv", sep = "" ))
  # save just the mean, with the variable names to an object
  assign(paste("colonydist+", var, "_k", i, sep = ""), 
       as_tibble(cbind(group = names(allevals2), t(allevals2))) %>%
         filter(group == "mean") %>% 
         select(-group) %>% 
         mutate(var = var, k = i) )
  rm(allevals2)
}
```