---
title: "5_GAMs"
author: "Gemma Clucas"
date: "3/24/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mgcv)
library(pROC)
library(tidyverse)
```

Read in my data
```{r}
data <- read.csv(file = "PresBackgroundLocationsWithEnvironmentalVariables.csv", stringsAsFactors = FALSE)

birds <- unique(data$Ptt)

names(data)

#birds <- "196698"
```


Run the GAMs varying the number of knots from 3 to 5. This holds out the real and background data for each Ptt in turn, and then calculates the mean AUC, sensitivity, and specificty from all 20 runs.

I can't seem to get the formula for the GAM to take anything but the actual variable name, so I am not able to run through each variable automatically at the moment.
```{r}
var <- "colonydist"
k <- c(3,4,5)
# k <- c(3)

for (i in k) {
  for (j in birds){
    TRAIN <- data[data$Ptt!=j,]  # use all groups except 1 in training data
    TEST<- data[data$Ptt==j,]  # keep the other group for testing data
    # run gam with one variable, k is number of knots
    GAM <- gam(pres ~ s(colonydist, k = i), data=TRAIN, bs="cs", family=binomial, select=TRUE, method='GCV.Cp')  
    TEST$GAM_pred <- as.numeric(predict(GAM, type="response", newdata=TEST)) # predict into the test data frame
    roc1 <- roc(TEST$pres, TEST$GAM_pred) # create a roc curve from the test data
    AUCEVAL <- as.numeric(roc1[9]) # get the auc value
    co1 <- pROC::coords(roc1, x = "best", best.method = "closest.topleft", ret=c("specificity","sensitivity")) #also get specificity and sensitivity
    # sometimes there is more than one "best" value
    # if just one, take those values
    # if(dim(co1)[1] == 1){
    #   spec1 <- as.numeric(co1[1])
    #   sen1 <- as.numeric(co1[2])
    #   # if not, take the first value for each - THIS IS OBVIOUSLY NOT IDEAL, BUT RUNS FOR NOW
    # } else {
    #   spec1 <- as.numeric(co1[1,1])
    #   sen1 <- as.numeric(co1[2,1])
    # }
    spec1 <- as.numeric(co1[1])
    sen1 <- as.numeric(co1[2])
    eva <- as.data.frame(rbind(AUCEVAL,spec1,sen1))
    names(eva) <- j
    eva
    if(exists("allevals")){
      allevals <- cbind(allevals, eva)
    }else {
      allevals <- eva
    }
  }
  # calculate mean AUC, sensitiviy, and specificity
  allevals$mean <- rowMeans(allevals)
  # save as a csv
  write.csv(allevals, file = paste("GAMs/", var, "_k", i, ".csv", sep = "" ))
  # save just the mean, with the variable names to an object
  assign(paste(var, i, sep = "_k"), 
       as_tibble(cbind(group = names(allevals), t(allevals))) %>%
         filter(group == "mean") %>% 
         select(-group) %>% 
         mutate(var = var, k = i) )
  rm(allevals)
}
```

Combine the outputs from all the different GAMs with just one covariate, to find the one with the best AUC. 
```{r}
  
dplyr::bind_rows(colonydist_k3, colonydist_k4, colonydist_k5) %>% 
                 # depth_k3, depth_k4, depth_k5,
                 # shelfdist_k3, shelfdist_k4, shelfdist_k5,
                 # slope_k3, slope_k4, slope_k5,
                 # SST_k3, SST_k4, SST_k5,
                 # Height_k3, Height_k4, Height_k5,
                 # chlorA_k3, chlorA_k4, chlorA_k5,
                 # NorthVelocity_k3, NorthVelocity_k4, NorthVelocity_k5,
                 # EastVelocity_k3, EastVelocity_k4, EastVelocity_k5) %>% 
  select(var, k, AUCEVAL, spec1, sen1) %>% 
  rename(covariate = var, knots = k,
         AUC = AUCEVAL, Sensitivity = sen1, Specificity = spec1) %>% 
  arrange(desc(AUC))
```

So colonydist is the best predictor. Run models with colonydist plus the other covariates next.
Also try using different methods for "best" threshold determination.