---
title: "Crawl PTT 196700"
author: "Gemma Clucas"
date: "6/29/2020"
output: github_document
---
  
```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(crawl)   #to fit Kalman filter models
library(trip)    #to prepare GPS data
library(maptools)
library(sp)
library(rgdal)
library(raster)
library(ggplot2)
library(knitr)
library(spdplyr)
options(scipen=999)
```

For an annotated version of this, look at ```1_Initial_Processing_with_Crawl.Rmd```

## Load data

This is the data that Claudia sent over on 30th June 2020. 

```{r}
raw <- read.csv("raw_data/satellite_tracking_data_30_06_20.csv")
```

## Prepare data for analysis

**Format times**  
  
```{r}
raw$Time <- mdy_hm(raw$Date, tz = "UTC")
raw$Time_since <- as.numeric(difftime(raw$Time, min(raw$Time), units="hours"))
raw <- raw[!duplicated(raw), ]
raw <- raw %>% 
  rename(Argos_loc_class = Loc.Class, LON = Lon1, LAT = Lat1)

clean <- raw %>% 
  dplyr::select(Ptt, Time, Time_since, LON, LAT, Argos_loc_class, Uplink)


```

## Create a map that we can plot the fixes onto later
```{r}
Seamask<-readOGR("Seamask.shp")
SSI <- crop(Seamask, c(450000, 750000, -600000, -100000))


#Re-project to Lambert Azimuthal Equal Area
SSI_laea<-spTransform(SSI, CRS=CRS("+proj=laea +lon_0=-26 +lat_0=-58 +units=m"))

# convert to dataframe for use with ggplot2
SSI_laea@data$id = rownames(SSI_laea@data)
SSI_laea.points = fortify(SSI_laea, region="id")
SSI_laea.df = plyr::join(SSI_laea.points, SSI_laea@data, by="id")

# filter out only the polygons for the islands
SSI_laea.df <- SSI_laea.df %>% filter(hole == TRUE)


SSI_laea.df %>% 
  ggplot(aes(x = long, y = lat, group = group)) + 
  geom_polygon(fill="grey") +
  geom_path(color="grey") +
  coord_equal() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "aliceblue"))
```


## Create functions for the different steps

```{r}
# define the function for selecting the data for one penguin and applying to McConnel speed filter
filter_points <-  function(penguin) {
  # select the data for the penguin
  x1 <- clean %>% dplyr::filter(Ptt == penguin)
  # order error classes
  x1$Argos_loc_class <- factor(x1$Argos_loc_class,  
                             levels=c("3","2","1","0", "A","B"))
  # adjust duplicate times
  x1$Time <- adjust.duplicateTimes(x1$Time, x1$Ptt)
  # select just the latitude, longitude, time and id columns
  x2 <- x1 %>% dplyr::select(LAT, LON, Time, Ptt)
  # make it spatial
  coordinates(x2) <- c("LON","LAT")
  # apply McConnel speed filter
  x2$Pass_speed <- x2 %>% 
    trip(., TORnames = c("Time","Ptt")) %>% 
    speedfilter(., max.speed = 8) 
  # filter the original data for the points that pass the speed filter
  x1 <- x1 %>% dplyr::filter(x2$Pass_speed == TRUE)
  # make the points spatial
  coordinates(x1) <- ~LON + LAT
  # give it a projection
  proj4string(x1) <- CRS("+proj=longlat +ellps=WGS84")
  # transform to LAEA centered around the south sandwich islands
  x1 <- spTransform(x1, CRS = CRS("+proj=laea +lon_0=-26 +lat_0=-58 +units=m"))
}


# define a function for plotting the cleaned data to check that it looks ok
plot_cleaned_points <- function(x) {
  ggplot() + 
    geom_polygon(data = SSI_laea.df, aes(x = long, y = lat, group = group), fill="grey80") +
    geom_path(data = SSI_laea.df, aes(x = long, y = lat, group = group), color="grey80") +
    coord_equal() +
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = "aliceblue")) +
    geom_point(data = x, aes(x = LON, y = LAT), colour = "orange")
}  

# define the prior for crawl
prior <-  function(p) { 
    dnorm(p[1], log(250), 0.2 , log = TRUE) +     # prior for 3
      dnorm(p[2], log(500), 0.2 , log = TRUE) +   # prior for 2
      dnorm(p[3], log(1500), 0.2, log = TRUE) +   # prior for 1
      dnorm(p[4], log(2500), 0.4 , log = TRUE) +  # prior for 0
      dnorm(p[5], log(2500), 0.4 , log = TRUE) +  # prior for A
      dnorm(p[6], log(2500), 0.4 , log = TRUE) +  # prior for B
      # skip p[7] as we won't provide a prior for sigma
      dnorm(p[8], -4, 2, log = TRUE)              # prior for beta
}


# define a function for crawling
crawl <- function(x1) {
  if(exists("fit1")){rm(fit1)}
  initial = list(a = c(coordinates(x1)[1,1], 0,
                     coordinates(x1)[1,2], 0),
               P = diag(c(10 ^ 2, 10 ^ 2, 10 ^ 2, 10 ^ 2)))
  fit1 <- crwMLE( 
    mov.model = ~1, 
    err.model=list(x=~Argos_loc_class-1), 
    drift=T, 
    data=x1, 
    Time.name="Time_since",  
    initial.state=initial, 
    prior=prior, 
    control=list(trace=1, REPORT=1)) 
}

# function for creating times where points will be predicted
predict_times <- function(x) {
  seq(min(x$Time_since), max(x$Time_since), 1/12)
}

# define a function for predicting points every 5 minutes, based on the crawled data
predict_points <- function(x, y) {
  crwPredict(object.crwFit = x, predTime = y, speedEst = TRUE, flat=TRUE)
}
  

# define a function for plotting the results of the crawled tracks against the original points
plot_crawled_track <- function() {
  # get the original data
  x3 <- clean %>% dplyr::filter(Ptt == penguin)
  # project original data to LAEA
  coordinates(x3) <- ~LON + LAT 
  proj4string(x3) <- CRS("+proj=longlat +ellps=WGS84")
  x3 <- spTransform(x3, CRS = CRS("+proj=laea +lon_0=-26 +lat_0=-58 +units=m"))
  # get just the coordinates from x3 and make a dataframe for plotting
  x3 <- coordinates(x3) %>% as.data.frame()
  # define the colour scheme
  colors <- c("Raw points" = "orangered", "McConnel speed filtered points" = "orange")
  # plot
  ggplot() + 
    geom_polygon(data = SSI_laea.df, aes(x = long, y = lat, group = group), fill="grey50") +
    geom_path(data = SSI_laea.df, aes(x = long, y = lat, group = group), color="grey50") +
    coord_equal() +
    geom_point(data = x3, aes(x = LON, y = LAT, colour = "Raw points") ) +
    geom_point(data = x1.df, aes(x = LON, y = LAT, colour = "McConnel speed filtered points")) +
    geom_path(data = predObj, aes(x = mu.x, y = mu.y)) +
    scale_color_manual(values = colors) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = "aliceblue"),
          legend.title = element_blank()) 
}

# zoom in on Saunders Island only
plot_crawled_track_Saunders <- function() {
  # get the original data
  x3 <- clean %>% dplyr::filter(Ptt == penguin)
  # project original data to LAEA
  coordinates(x3) <- ~LON + LAT 
  proj4string(x3) <- CRS("+proj=longlat +ellps=WGS84")
  x3 <- spTransform(x3, CRS = CRS("+proj=laea +lon_0=-26 +lat_0=-58 +units=m"))
  # get just the coordinates from x3 and make a dataframe for plotting
  x3 <- coordinates(x3) %>% as.data.frame()
  # define the colour scheme
  colors <- c("Raw points" = "orangered", "McConnel speed filtered points" = "orange")
  # plot
  ggplot() + 
    geom_polygon(data = SSI_laea.df, aes(x = long, y = lat, group = group), fill="grey50") +
    geom_path(data = SSI_laea.df, aes(x = long, y = lat, group = group), color="grey50") +
    coord_equal() +
    geom_point(data = x3, aes(x = LON, y = LAT, colour = "Raw points") ) +
    geom_point(data = x1.df, aes(x = LON, y = LAT, colour = "McConnel speed filtered points")) +
    geom_path(data = predObj, aes(x = mu.x, y = mu.y)) +
    scale_color_manual(values = colors) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = "aliceblue"),
          legend.title = element_blank()) +
    coord_cartesian(xlim = c(-50000, 75000),
                    ylim = c(-50000, 75000))

}


```


```{r}
penguin <- "196700"
x1 <- filter_points(penguin)
# convert to dataframe for plotting
x1.df <- data.frame(x1)
plot_cleaned_points(x1.df)

```


Get rid of that weird fix.
```{r}
x1.df <- x1.df %>% dplyr::filter(LON > -500000)
plot_cleaned_points(x1.df)

# then make x1.df spatial 
coordinates(x1.df) <- ~LON + LAT
# tell it it's already projected in LAEA centered around the south sandwich islands
proj4string(x1.df) <- CRS("+proj=laea +lon_0=-26 +lat_0=-58 +units=m")
# transform to LAEA centered around the south sandwich islands
x1.df <- spTransform(x1.df, CRS = CRS("+proj=laea +lon_0=-26 +lat_0=-58 +units=m"))

# rename back to x1 so that I can continue with the code below
x1 <- x1.df

# then recreate x1.df as a normal data frame
x1.df <- data.frame(x1)

```


Crawl (ran many times).

```{r crawl_196700, cache = TRUE}
if(exists("fit1")){rm(fit1)} 
fit1 <- crawl(x1)
print(fit1)
```

This was the best I could get, but the tracks look ok below

```{r}
predTime <- predict_times(x1)
predObj <- predict_points(fit1, predTime)

plot_crawled_track()
plot_crawled_track_Saunders()
```
**A lot of points seem to have been filtered by the speed filter - should it be more lenient?**


Write to CSV:
```{r}
write.csv(predObj, paste0("predicted_tracks/", penguin, "_track.csv", sep = ""), row.names = FALSE)
```


